Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-09-30 Manuscript/HI_Model_ms.Rmd
updated manuscript files

## Abstract
The State of Hawai`i passed legislation in 2018 setting a goal to be carbon neutral by 2045. Meeting this goal will partly depend on carbon sequestration by terrestrial ecosystems, yet the future direction and magnitude of the land carbon sink in the Hawaiian Islands is highly uncertain. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under four unique scenarios over a 90-year timespan. Net ecosystem carbon balance declined under all four scenarios. Moving from a high to a low radiative forcing scenario reduced net ecosystem carbon loss by ~21%, and net carbon losses were reduced by a total of ~55% under the combined scenario of low radiative forcing and low rates of land‐use change. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to better constrainment of models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction
The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,

>>>>

## Abstract
The State of Hawai`i passed legislation to be carbon neutral by 2045, a goal that partly depends on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCPs 4.5 and 8.5) over a 90-year timespan from 2010-2100. Collectively, ecocystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing throughout the 90-year simulation period, but carbon sink strength capacity was greatest under the combination of low radiative forcing and low land-use change. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing, with high land-use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Reconciling this uncertainty in how net photosynthesis will respond to rising atmospheric CO~2~ will be essential to constrain models used to evaluate the effectiveness of ecosystem‐based climate mitigation strategies.

## Introduction
Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis [Friedlingstein_global_2019].

The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances.
